---
title: "Figure"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r setup}
library(tidyverse)
library(ggpubr)
```

```{r color}
disease_list = c("normal", "COPD", "IPF", "COVID-19", "cancer")
disease_cols = c("normal" = "#780001",
                 "COPD" = "#C11221",
                 "IPF" = "#FEF0D5",
                 "COVID-19" = "#002F49",
                 "cancer" = "#669BBB")
```

## Figure 3 - Overview of datasets
### Figure 3A - Disease
```{r Figure 3A}
metadata <- read_csv("Figure/Fig3/metadata.csv")
dataset <- metadata %>%
  select(dataset, disease) %>%
  unique() %>%
  group_by(disease) %>% 
  summarise(count = n())

patient <- metadata %>%
  select(sampleID, disease) %>%
  unique() %>%
  group_by(disease) %>% 
  summarise(count = n())
```

### Figure 3B - Metadata
```{r Figure 3B sex barplot}
metadata <- readr::read_csv("Figure/metadata.csv")
sex <- metadata %>% 
  filter(sex != "unknown") %>%
  select(sampleID, disease, sex) %>% 
  group_by(sex, disease) %>% 
  summarise(count = n())
sex$sex <- factor(sex$sex, levels = c("Male", "Female"))
sex$disease <- factor(sex$disease, levels = disease_list)
ggplot(sex, aes(x = sex, y = count, fill = disease)) +
  geom_col(position = "stack", width = 0.7) +
  theme_classic() +
  scale_fill_manual(values = disease_cols) +
  labs(x = "Gender", y = "Number of samples", fill = "Disease") +
  theme(panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        legend.position = "none")
```

```{r Figure 3B disease legend}
sex_barplot <- ggplot(sex, aes(x = sex, y = count, fill = disease)) +
  geom_col(position = "stack", width = 0.7) +
  theme_classic() +
  scale_fill_manual(values = disease_cols) +
  labs(x = "Gender", y = "Number of samples", fill = "Disease") +
  theme(panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        legend.position = "right")
legend <- get_legend(sex_barplot)
as_ggplot(legend)
```

```{r Figure 3B sex NA}
metadata <- readr::read_csv("Figure/metadata.csv")
wd <- c("unknown" = "unknown", "Male" = "known", "Female" = "known")
sex <- metadata %>% 
  select(sampleID, disease, sex)
sex$wd <- factor(wd[sex$sex], levels = c("unknown", "known"))
sex <- sex %>%
  group_by(wd) %>% 
  summarise(count = n())

ggplot(sex, aes(x = 1, y = count, fill = wd)) +
  geom_col(position = "fill", width = 0.7) +
  theme_classic() +
  scale_fill_manual(values = c("#e8e8e8", "#555555")) +
  labs(x = "", y = "Proportion of samples", fill = "Data availability") +
  theme(axis.line = element_blank(),
        axis.ticks.x = element_blank(),
        axis.text.x = element_blank(),
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        legend.position = "none")
```

```{r Figure 3B data availability legend}
sex_barplot <- ggplot(sex, aes(x = 1, y = count, fill = wd)) +
  geom_col(position = "fill", width = 0.7) +
  theme_classic() +
  scale_fill_manual(values = c("#e8e8e8", "#555555")) +
  labs(x = "", y = "Proportion of samples", fill = "Data availability") +
  theme(axis.line = element_blank(),
        axis.ticks.x = element_blank(),
        axis.text.x = element_blank(),
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        legend.position = "right")
legend <- get_legend(sex_barplot)
as_ggplot(legend)
```

```{r Figure 3B smoking barplot}
smoking <- metadata %>% 
  filter(smoking_status != "unknown") %>%
  select(sampleID, disease, smoking_status) %>% 
  group_by(smoking_status, disease) %>% 
  summarise(count = n())
smoking$smoking_status <- factor(smoking$smoking_status, levels = c("Yes", "No", "unknown"))
smoking$disease <- factor(smoking$disease, levels = disease_list)
ggplot(smoking, aes(x = smoking_status, y = count, fill = disease)) +
  geom_col(position = "stack", width = 0.7) +
  theme_classic() +
  scale_fill_manual(values = disease_cols) +
  labs(x = "Smoking status", y = "Number of samples", fill = "Disease") +
  theme(panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        legend.position = "none")
```

```{r Figure 3B smoking NA}
wd <- c("unknown" = "unknown", "Yes" = "known", "No" = "known")
smoking <- metadata %>% 
  select(sampleID, disease, smoking_status)
smoking$wd <- factor(wd[smoking$smoking_status], levels = c("unknown", "known"))
smoking <- smoking %>%
  group_by(wd) %>% 
  summarise(count = n())

ggplot(smoking, aes(x = 1, y = count, fill = wd)) +
  geom_col(position = "fill", width = 0.7) +
  theme_classic() +
  scale_fill_manual(values = c("#e8e8e8", "#555555")) +
  labs(x = "", y = "Proportion of samples", fill = "Data availability") +
  theme(axis.line = element_blank(),
        axis.ticks.x = element_blank(),
        axis.text.x = element_blank(),
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        legend.position = "none")
```

```{r Figure 3B age barplot}
lage <- c("30-35" = "30-40",
          "40-45" = "40-50",
          "45-50" = "40-50",
          "50-55" = "50-60",
          "55-60" = "50-60",
          "60-65" = "60-70",
          "65-70" = "60-70",
          "70-75" = "70-80",
          "75-80" = "70-80",
          "80-85" = "80-90",
          "85-90" = "80-90")
age <- metadata %>% 
  filter(age != "unknown") %>%
  select(sampleID, disease, age)
age$age <- lage[age$age]
age <- age %>%
  group_by(age, disease) %>% 
  summarise(count = n())
age$disease <- factor(age$disease, levels = disease_list)

ggplot(age, aes(x = age, y = count, fill = disease)) +
  geom_col(position = "stack", width = 0.7) +
  theme_classic() +
  scale_fill_manual(values = disease_cols) +
  labs(x = "Age", y = "Number of samples", fill = "Disease") +
  theme(panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        legend.position = "none")
```

```{r Figure 3B age NA}
wd <- c("unknown" = "unknown")
age <- metadata %>% 
  select(sampleID, disease, age)
age$wd <- wd[age$age]
age$wd <- ifelse(is.na(age$wd), "known", "unknown")
age$wd <- factor(age$wd, levels = c("unknown", "known"))
age <- age %>%
  group_by(wd) %>% 
  summarise(count = n())

ggplot(age, aes(x = 1, y = count, fill = wd)) +
  geom_col(position = "fill", width = 0.7) +
  theme_classic() +
  scale_fill_manual(values = c("#e8e8e8", "#555555")) +
  labs(x = "", y = "Proportion of samples", fill = "Data availability") +
  theme(axis.line = element_blank(),
        axis.ticks.x = element_blank(),
        axis.text.x = element_blank(),
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        legend.position = "none")
```

## Figure 5 - Stroma/Endo cell counts
```{r Figure 5A stroma/endo barplot}
freq = data.frame(disease = disease_list,
                  Stromal = c(25217, 12654, 36665, 24081, 8021),
                  Endothelial = c(48166, 17805, 13190, 11199, 6830))
freq = pivot_longer(freq, cols = c(Stromal, Endothelial), 
                    names_to = "celltype", values_to = "freq")
freq$celltype = factor(freq$celltype, levels = c("Stromal", "Endothelial"))
freq$disease = factor(freq$disease, levels = disease_list)
ggplot(freq, aes(x = celltype, y = freq, fill = disease)) +
  geom_col(position = "fill", width = 0.7) +
  scale_fill_manual(values = disease_cols) +
  labs(x = "", y = "Frequency", fill = "Disease") +
  theme_classic()
```

```{r Figure 5B stroma frequency boxplot}
library(readr)
freq <- readr::read_csv("Figure/Fig5/cellcounts.csv")
freq <- freq %>% filter(Stroma >= 100)
freq_disease <- freq[, 6]
freq <- freq[, 2:5]
freq <- cbind(freq/rowSums(freq), freq_disease)
freq$disease <- factor(freq$disease, levels = disease_list)
freq <- select(freq, Stroma, disease)
ggplot(freq, aes(x = disease, y = Stroma, fill = disease)) +
  geom_boxplot(outlier.shape = NA) +
  geom_jitter(width = 0.1, size = 1.5) +
  stat_compare_means(method = "wilcox.test", ref.group = "normal",
                     label = "p.format", label.y = 0.8,
                     size = 3) +
  theme_classic() +
  scale_fill_manual(values = disease_cols) +
  labs(x = "Disease", y = "Frequency of stromal cells") +
  theme(panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        legend.position = "none")
```

```{r Figure 5B endo frequency boxplot}
freq <- readr::read_csv("Figure/Fig5/cellcounts.csv")
freq <- freq %>% filter(Endothelial >= 80)
freq_disease <- freq[, 6]
freq <- freq[, 2:5]
freq <- cbind(freq/rowSums(freq), freq_disease)
freq$disease <- factor(freq$disease, levels = disease_list)
freq <- select(freq, Endothelial, disease)
ggplot(freq, aes(x = disease, y = Endothelial, fill = disease)) +
  geom_boxplot(outlier.shape = NA) +
  geom_jitter(width = 0.15, size = 1.5) +
  stat_compare_means(method = "wilcox.test", ref.group = "normal",
                     label = "p.format", label.y = 0.32,
                     size = 3) +
  theme_classic() +
  scale_fill_manual(values = disease_cols) +
  labs(x = "Disease", y = "Frequency of endothelial cells") +
  theme(panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        legend.position = "none")
```

## Figure 10 - Stroma DEG & GSEA
```{r Figure 10B alveolar gsea}
alv_gsea <- readr::read_csv("3.Stroma/alv_gsea.csv")
hallmark <- gsub("HALLMARK", "", gsub("_", " ", alv_gsea$source))
hallmark <- stringr::str_to_title(hallmark)
alv_gsea$source <- hallmark

ggplot(alv_gsea[c(1:5), ], aes(y = reorder(source, norm), x = norm)) +
  geom_col(width = 0.8, fill = "#C62828") +
  theme_classic() +
  labs(x = "Normalized Enrichment Score", y = "Hallmark Entries") +
  theme(panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        legend.position = "none")
```

```{r Figure 10D myofibro gsea}
myo_gsea <- readr::read_csv("3.Stroma/myo_gsea.csv")
hallmark <- gsub("HALLMARK", "", gsub("_", " ", myo_gsea$source))
hallmark <- stringr::str_to_title(hallmark)
myo_gsea$source <- hallmark
myo_gsea$positive <- myo_gsea$norm > 0

ggplot(myo_gsea[c(1:5, 46:49),], 
       aes(y = reorder(source, norm), x = norm, fill = positive)) +
  geom_col(width = 0.8) +
  theme_classic() +
  labs(x = "Normalized Enrichment Score", y = "Hallmark Entries") +
  scale_fill_manual(values = c("#3F51B5", "#29B6F6")) +
  theme(panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        legend.position = "none")
```

## Figure 11 - Stroma & Disease
### Figure 11A - Stroma Ro/e
```{r Figure 11A stroma Ro/e}
library(ComplexHeatmap)
library(RColorBrewer)
stroma_roe <- readr::read_csv("3.Stroma/stroma_ROE.csv")
stroma_roe <- stroma_roe %>% 
  column_to_rownames("minor_cell_type") %>%
  select(disease_list)
stroma_roe <- as.matrix(log(stroma_roe))
pheatmap(stroma_roe, 
         cellwidth = 20, cellheight = 20,
         color = colorRampPalette(rev(brewer.pal(n = 11, name = "RdBu")))(100),
         cluster_cols = FALSE, cluster_rows= TRUE, clustering_distance_rows = "pearson",
         treeheight_col = 0, treeheight_row = 15,
         cutree_row = 3,
         border_color = "black")
```
### Figure 11B - Frequency of stroma minor cell types
```{r Figure 11B stroma minor cell type}
library(readr)
freq <- readr::read_csv("3.Stroma/stroma_minor_cell_type.csv")
freq$all <- rowSums(freq[, -c(1,16)])
freq <- freq %>% filter(all >= 100) %>% select(-all)
freq_disease <- freq[, 16]
freq <- freq[, 2:15]
freq <- cbind(freq/rowSums(freq), freq_disease)
freq$disease <- factor(freq$disease, levels = disease_list)
ggplot(freq, aes(x = disease, y = `Myofibro. CCL19+`, fill = disease)) +
  geom_boxplot(outlier.shape = NA) +
  geom_jitter(width = 0.1, size = 1) +
  stat_compare_means(method = "wilcox.test",
                     comparisons = list(c(5,1), c(5,2), c(5,3)),
                     label = "p.signif", size = 3, tip.length = 0) +
  theme_classic() +
  scale_fill_manual(values = disease_cols) +
  labs(title = "Myofibro. CCL19+", x = "Disease", y = "Proportions of all stromal cells") +
  theme(panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        legend.position = "none")

ggplot(freq, aes(x = disease, y = `Myofibro. CCL19-`, fill = disease)) +
  geom_boxplot(outlier.shape = NA) +
  geom_jitter(width = 0.1, size = 1) +
  stat_compare_means(method = "wilcox.test",
                     comparisons = list(c(4,1), c(4,2), c(5,1), c(5,2), c(5,3)),
                     label = "p.signif", size = 3, tip.length = 0) +
  theme_classic() +
  scale_fill_manual(values = disease_cols) +
  labs(title = "Myofibro. CCL19-", x = "Disease", y = "Proportions of all stromal cells") +
  theme(panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        legend.position = "none")

ggplot(freq, aes(x = disease, y = `SMCs`, fill = disease)) +
  geom_boxplot(outlier.shape = NA) +
  geom_jitter(width = 0.1, size = 1) +
  stat_compare_means(method = "wilcox.test",
                     comparisons = list(c(2,4), c(2,5), c(3,4), c(3,5)),
                     label = "p.signif", size = 3, tip.length = 0) +
  theme_classic() +
  scale_fill_manual(values = disease_cols) +
  labs(title = "SMCs", x = "Disease", y = "Proportions of all stromal cells") +
  theme(panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        legend.position = "none")

ggplot(freq, aes(x = disease, y = `Adv.fibro.`, fill = disease)) +
  geom_boxplot(outlier.shape = NA) +
  geom_jitter(width = 0.1, size = 1) +
  stat_compare_means(method = "wilcox.test",
                     comparisons = list(c(1,4), c(1,5)),
                     label = "p.format", size = 3, tip.length = 0) +
  theme_classic() +
  scale_fill_manual(values = disease_cols) +
  labs(title = "Adv.fibro.", x = "Disease", y = "Proportions of all stromal cells") +
  theme(panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        legend.position = "none")
```

## Figure 13 - Endo clustering
```{r Figure 13E endo aucell}
pathway_score <- read.csv("4.Endo/aucell_pathway.csv")
pathway_score <- pathway_score %>%
  select(-X) %>%
  group_by(minor_cell_type) %>%
  summarise_each(funs(mean))
pathway_score <- pathway_score %>% 
  column_to_rownames("minor_cell_type") %>% 
  t() %>% 
  as.data.frame()
pathway_score <- pathway_score[c(3,2,5,1,4),]
rownames(pathway_score) <- c("Leukocyte adhesion to vascular endo.",
                             "Antigen processing and presentation",
                             "Interferon signaling",
                             "VEGF pathway",
                             "Collagen formation")

library("pheatmap")
pheatmap(pathway_score, 
         cellwidth = 20, cellheight = 20,
         color = colorRampPalette(rev(brewer.pal(n = 11, name = "RdBu")))(100),
         scale = "row",
         cluster_cols = FALSE, cluster_rows= FALSE,
         treeheight_col = 0, treeheight_row = 0,
         border_color = "black")
```

## Figure 14 - Endo & Disease
### Figure 14A - Endo Ro/e
```{r Figure 14A endo Ro/e}
library(ComplexHeatmap)
library(RColorBrewer)
endo_roe <- readr::read_csv("4.Endo/endo_ROE.csv")
endo_roe <- endo_roe %>% 
  column_to_rownames("minor_cell_type") %>%
  select(disease_list)
endo_roe <- as.matrix(log(endo_roe))
pheatmap(endo_roe, 
         cellwidth = 20, cellheight = 20,
         color = colorRampPalette(rev(brewer.pal(n = 11, name = "RdBu")))(100),
         breaks = seq(-1, 1, length.out = 100),
         cluster_cols = FALSE, cluster_rows= TRUE, clustering_distance_rows = "pearson",
         treeheight_col = 0, treeheight_row = 15,
         cutree_rows = 3,
         border_color = "black")
```

### Figure 14B - Frequency of endo minor cell types
```{r Figure 14B endo all minor cell type}
library(readr)
freq <- readr::read_csv("4.Endo/endo_minor_cell_type.csv")
freq$all <- rowSums(freq[, -c(1,11)])
freq <- freq %>% filter(all >= 80) %>% select(-all)
freq_disease <- freq[, 11]
freq <- freq[, 2:10]
freq <- cbind(freq/rowSums(freq), freq_disease)
freq$disease <- factor(freq$disease, levels = disease_list)

freq <- freq %>% 
  pivot_longer(cols = !disease, names_to = "minor_cell_type", values_to = "freq")
ggplot(freq, aes(x = disease, y = freq, fill = disease)) +
  geom_boxplot(outlier.shape = NA) +
  geom_jitter(width = 0.1, size = 1) +
  stat_compare_means(aes(label = sprintf("p = %4.3f", as.numeric(..p.format..))),
                     method = "wilcox.test", ref.group = ".all.", size = 3) +
  theme_classic() +
  facet_wrap("minor_cell_type", scale = "free_y", ncol = 2) +
  scale_fill_manual(values = disease_cols) +
  labs(x = "Disease", y = "Proportions of all endothelial cells") +
  theme(panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        legend.position = "none")
```

```{r Figure 14B endo selected minor cell type}
freq <- readr::read_csv("4.Endo/endo_minor_cell_type.csv")
freq$all <- rowSums(freq[, -c(1,11)])
freq <- freq %>% filter(all >= 80) %>% select(-all)
freq_disease <- freq[, 11]
freq <- freq[, 2:10]
freq <- cbind(freq/rowSums(freq), freq_disease)
freq$disease <- factor(freq$disease, levels = disease_list)
ggplot(freq, aes(x = disease, y = `endo. PLVAP`, fill = disease)) +
  geom_boxplot(outlier.shape = NA) +
  geom_jitter(width = 0.1, size = 1) +
  stat_compare_means(method = "wilcox.test",
                     comparisons = list(c(3,1), c(3,2), c(3,4), c(5,1), c(5,2), c(5,4)),
                     label = "p.signif", size = 3, tip.length = 0) +
  theme_classic() +
  scale_fill_manual(values = disease_cols) +
  labs(title = "endo. PLVAP", x = "Disease", y = "Proportions of all endothelial cells") +
  theme(panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        legend.position = "none")

ggplot(freq, aes(x = disease, y = `aCap` + `gCap`, fill = disease)) +
  geom_boxplot(outlier.shape = NA) +
  geom_jitter(width = 0.1, size = 1) +
  stat_compare_means(method = "wilcox.test",
                     comparisons = list(c(3,1), c(3,2), c(3,4), c(3,5)),
                     label = "p.signif", size = 3, tip.length = 0) +
  theme_classic() +
  scale_fill_manual(values = disease_cols) +
  labs(title = "aCap + gCap", x = "Disease", y = "Proportions of all endothelial cells") +
  theme(panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        legend.position = "none")

ggplot(freq, aes(x = disease, y = `endo. CXCL10`, fill = disease)) +
  geom_boxplot(outlier.shape = NA) +
  geom_jitter(width = 0.1, size = 1) +
  stat_compare_means(method = "wilcox.test",
                     comparisons = list(c(3,1), c(3,2), c(3,4), c(3,5)),
                     label = "p.signif", size = 3, tip.length = 0) +
  theme_classic() +
  scale_fill_manual(values = disease_cols) +
  labs(title = "endo. CXCL10", x = "Disease", y = "Proportions of all endothelial cells") +
  theme(panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        legend.position = "none")

ggplot(freq, aes(x = disease, y = `sys.Vein`, fill = disease)) +
  geom_boxplot(outlier.shape = NA) +
  geom_jitter(width = 0.1, size = 1) +
  stat_compare_means(method = "wilcox.test",
                     comparisons = list(c(3,1), c(3,2), c(3,4), c(3,5)),
                     label = "p.signif", size = 3, tip.length = 0) +
  theme_classic() +
  scale_fill_manual(values = disease_cols) +
  labs(title = "sys.Vein", x = "Disease", y = "Proportions of all endothelial cells") +
  theme(panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        legend.position = "none")
```

## Figure 16 - Stroma label transfer
### Figure 16C - Performance evaluation  
```{r Figure 16C stroma uncertainty}
uncert <- read_csv("5.LabelTransfer/stroma_uncertainty.csv")
uncert$minor_cell_type <- factor(uncert$minor_cell_type)
uncert$minor_cell_type <- factor(uncert$minor_cell_type,
                                 levels=rev(levels(uncert$minor_cell_type)))
celltype_cols = rev(c("#673AB7", "#C62828", "#F44336", "#006064",
                      "#607D8B", "#29B6F6", "#3F51B5", "#009688", 
                      "#4CAF50", "#8BC34A", "#9E9E9E", "#FFEB3B", 
                      "#FF9800", "#795548"))
ggplot(uncert, aes(x = uncertainty, y = minor_cell_type, 
                   fill = minor_cell_type, group = minor_cell_type)) +
  geom_violin() + 
  labs(x = "Uncertainty") +
  scale_fill_manual(values = celltype_cols) +
  theme_classic() +
  theme(legend.position = "none") + 
  theme(axis.ticks.y = element_blank(),
        axis.text.y = element_blank(),
        axis.title.y = element_blank(),
        axis.line.y = element_blank())
```

```{r Figure 16C entropy & heatmap}
ltcelltype <- read_csv("5.LabelTransfer/stroma_celltype.csv")
ltcelltype <- ltcelltype %>%
  group_by(leiden, minor_cell_type) %>%
  summarise(counts = n()) %>%
  ungroup() %>%
  pivot_wider(names_from = minor_cell_type, values_from = counts) %>%
  column_to_rownames("leiden")
ltcelltype <- ltcelltype / rowSums(ltcelltype, na.rm = TRUE)
entropy <- apply(ltcelltype, 1, function(z)
  -sum(z * log10(z), na.rm = TRUE) / log10(sum(!is.na(z))))
entropy <- as.data.frame(entropy) %>% rownames_to_column("leiden")
entropy$leiden <- as.numeric(entropy$leiden)

ggplot(entropy, aes(x = leiden, y = entropy)) +
  geom_col(width = 0.8, fill = "#A50F15") + 
  labs(y = "Entropy") +
  theme_classic() +
  theme(axis.ticks.x = element_blank(),
        axis.text.x = element_blank(),
        axis.title.x = element_blank(),
        axis.line.x = element_blank())

ltcelltype[is.na(ltcelltype)] <- 0
pheatmap(as.matrix(t(ltcelltype > 1/7)),
         cellheight = 20, cellwidth = 20,
         color = brewer.pal(n = 11, name = "Greys"),
         cluster_cols = FALSE, cluster_rows= FALSE, 
         border_color = "black")
```

## Figure 17 - Endo label transfer
### Figure 17C - Performance evaluation 
```{r Figure 17C endo uncertainty}
uncert <- read_csv("5.LabelTransfer/endo_uncertainty.csv")
uncert$minor_cell_type <- factor(uncert$minor_cell_type)
uncert$minor_cell_type <- factor(uncert$minor_cell_type,
                                 levels=rev(levels(uncert$minor_cell_type)))
celltype_cols = rev(c("#673AB7", "#C62828", "#006064", 
                 "#29B6F6", "#795548", "#FF9800", 
                 "#607D8B", "#4CAF50", "#009688"))
ggplot(uncert, aes(x = uncertainty, y = minor_cell_type, 
                   fill = minor_cell_type, group = minor_cell_type)) +
  geom_violin() + 
  labs(x = "Uncertainty") +
  scale_fill_manual(values = celltype_cols) +
  theme_classic() +
  theme(legend.position = "none") + 
  theme(axis.ticks.y = element_blank(),
        axis.text.y = element_blank(),
        axis.title.y = element_blank(),
        axis.line.y = element_blank())
```

```{r Figure 17C endo entropy & heatmap}
ltcelltype <- read_csv("5.LabelTransfer/endo_celltype.csv")
ltcelltype <- ltcelltype %>%
  group_by(leiden, minor_cell_type) %>%
  summarise(counts = n()) %>%
  ungroup() %>%
  pivot_wider(names_from = minor_cell_type, values_from = counts) %>%
  column_to_rownames("leiden")
ltcelltype <- ltcelltype / rowSums(ltcelltype, na.rm = TRUE)
entropy <- apply(ltcelltype, 1, function(z)
  -sum(z * log10(z), na.rm = TRUE) / log10(sum(!is.na(z))))
entropy <- as.data.frame(entropy) %>% rownames_to_column("leiden")
entropy$leiden <- as.numeric(entropy$leiden)

ggplot(entropy, aes(x = leiden, y = entropy)) +
  geom_col(width = 0.8, fill = "#A50F15") + 
  labs(y = "Entropy") +
  theme_classic() +
  theme(axis.ticks.x = element_blank(),
        axis.text.x = element_blank(),
        axis.title.x = element_blank(),
        axis.line.x = element_blank())

ltcelltype[is.na(ltcelltype)] <- 0
pheatmap(as.matrix(t(ltcelltype > 2/9)),
         cellheight = 20, cellwidth = 20,
         color = brewer.pal(n = 11, name = "Greys"),
         cluster_cols = FALSE, cluster_rows= FALSE, 
         border_color = "black")
```

### Figure 18 - Frequency of transferred label
```{r Figure 18A stroma response frequency}
library(ggpubr)
ltcelltype <- read_csv("5.LabelTransfer/stroma_celltype.csv")
major_response <- unique(ltcelltype[, c("sampleID", "Major Response")])
major_response <- major_response %>% column_to_rownames("sampleID")
ltcelltype <- ltcelltype %>% 
  group_by(sampleID, minor_cell_type) %>%
  summarise(counts = n()) %>%
  ungroup() %>%
  pivot_wider(names_from = minor_cell_type, values_from = counts) %>%
  column_to_rownames("sampleID")
ltcelltype <- ltcelltype / rowSums(ltcelltype, na.rm = TRUE)
ltcelltype[is.na(ltcelltype)] <- 0
ltcelltype <- ltcelltype%>%
  rownames_to_column("sampleID") %>%
  pivot_longer(!sampleID, names_to = "minor_cell_type", values_to = "frequency")
ltcelltype$major_response <- major_response[ltcelltype$sampleID, ]

freq <- readr::read_csv("3.Stroma/stroma_minor_cell_type.csv")
freq$all <- rowSums(freq[, -c(1,16)])
freq <- freq %>% filter(all >= 100) %>% select(-all)
freq_disease <- freq[, 16]
freq <- freq[, 2:15]
freq <- cbind(freq/rowSums(freq), freq_disease)

freq <- freq %>% 
  filter(disease == "cancer") %>%
  rownames_to_column("sampleID") %>%
  pivot_longer(!c(sampleID, disease), names_to = "minor_cell_type", values_to = "frequency")
colnames(freq) <- c("sampleID", "major_response", "minor_cell_type", "frequency")
ltcelltype <- rbind(ltcelltype, freq)
ltcelltype$major_response <- factor(ltcelltype$major_response,
                                    levels = c("MPR", "non-MPR", "cancer"))
ltcelltype <- ltcelltype %>% filter(minor_cell_type %in%
                                      c("Alv.fibro. CCL2+",
                                        "Alv.fibro. CCL2-",
                                        "fibro. APOD",
                                        "Myofibro. CCL19+",
                                        "Myofibro. CCL19-"))

ggplot(ltcelltype, aes(x = major_response, y = frequency, fill = major_response)) +
  geom_boxplot(outlier.size = 0.5) +
  stat_compare_means(method = "wilcox.test",
                     comparison = list(c(1,2), c(1,3), c(2,3)),
                     label = "p.format", hide.ns = TRUE,
                     tip.length = 0,
                     size = 3) +
  labs(x = "Stromal cell subtypes", y = "Proportion of stromal cells") +
  facet_wrap("minor_cell_type", scales = "free", nrow = 1) +
  scale_fill_manual(values = c("#DFEBD5", "#CFDAEC", "#BFBBBA")) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1)) +
  theme(legend.position = "none")
```

```{r Figure 18B endo response frequency}
library(ggpubr)
ltcelltype <- read_csv("5.LabelTransfer/endo_celltype.csv")
major_response <- unique(ltcelltype[, c("sampleID", "major_response")])
major_response <- major_response %>% column_to_rownames("sampleID")
ltcelltype <- ltcelltype %>% 
  group_by(sampleID, minor_cell_type) %>%
  summarise(counts = n()) %>%
  ungroup() %>%
  pivot_wider(names_from = minor_cell_type, values_from = counts) %>%
  column_to_rownames("sampleID")
ltcelltype <- ltcelltype / rowSums(ltcelltype, na.rm = TRUE)
ltcelltype[is.na(ltcelltype)] <- 0
ltcelltype <- ltcelltype%>%
  rownames_to_column("sampleID") %>%
  pivot_longer(!sampleID, names_to = "minor_cell_type", values_to = "frequency")
ltcelltype$major_response <- major_response[ltcelltype$sampleID, ]

freq <- readr::read_csv("4.Endo/endo_minor_cell_type.csv")
freq$all <- rowSums(freq[, -c(1,11)])
freq <- freq %>% filter(all >= 80) %>% select(-all)
freq_disease <- freq[, 11]
freq <- freq[, 2:10]
freq <- cbind(freq/rowSums(freq), freq_disease)

freq <- freq %>% 
  filter(disease == "cancer") %>%
  rownames_to_column("sampleID") %>%
  pivot_longer(!c(sampleID, disease), names_to = "minor_cell_type", values_to = "frequency")
colnames(freq) <- c("sampleID", "major_response", "minor_cell_type", "frequency")
ltcelltype <- rbind(ltcelltype, freq)
ltcelltype$major_response <- factor(ltcelltype$major_response,
                                    levels = c("MPR", "non-MPR", "cancer"))
ltcelltype <- ltcelltype %>% filter(minor_cell_type %in%
                                      c("aCap",
                                        "Artery",
                                        "gCap",
                                        "pul.Vein",
                                        "sys.Vein"))

ggplot(ltcelltype, aes(x = major_response, y = frequency, fill = major_response)) +
  geom_boxplot(outlier.size = 0.5) +
  stat_compare_means(method = "wilcox.test",
                     comparison = list(c(1,2), c(1,3), c(2,3)),
                     label = "p.format", hide.ns = TRUE,
                     tip.length = 0,
                     size = 3) +
  labs(x = "Endothelial cell subtypes", y = "Proportion of endothelial cells") +
  facet_wrap("minor_cell_type", scales = "free", nrow = 1) +
  scale_fill_manual(values = c("#DFEBD5", "#CFDAEC", "#BFBBBA")) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1)) +
  theme(legend.position = "none")
```